Determining sequence-dependent DNA oligonucleotide hybridization and dehybridization mechanisms using coarse-grained molecular simulation, Markov state models, and infrared spectroscopy

### 2021-05-18 manuscript.tex
Andrei edits

%\newcommand*{\rood}[1]{{\color{red}{#1}}}
%\newcommand*{\bood}[1]{{\color{blue}{#1}}}
\newcommand*{\blauw}[1]{\textcolor{blue}{#1}}

%\newcommand*{\addref}[1]{\textcolor{red}{\{ADD REF: #1\}}}
%\newcommand*{\noter}[1]{\textcolor{red}{[[#1]]}}		% notes 
%\newcommand*{\noteb}[1]{\textcolor{blue}{[[#1]]}}		% notes 
%\newcommand*{\noter}[1]{}					% notes off

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Meta-data block
%% ---------------
%% Each author should be given as a separate \author command.
%%
%% Corresponding authors should have an e-mail given after the author
%% name as an \email command. Phone and fax numbers can be given
%% using \phone and \fax, respectively; this information is optional.
%%
%% The affiliation of authors is given after the authors; each
%% \affiliation command applies to all preceding authors not already
%% assigned an affiliation.
%%
%% The affiliation takes an option argument for the short name.  This
%% will typically be something like "University of Somewhere".
%%
%% The \altaffiliation macro should be used for new address, etc.
%% On the other hand, \alsoaffiliation is used on a per author basis
%% when authors are associated with multiple institutions.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\author{Michael S. Jones}
\affiliation{%
  Pritzker School of Molecular Engineering, %
  The University of Chicago, %
  929 East 57th Street, Chicago, Illinois 60637, United States%
}

@@

\author{Andrew L. Ferguson}
\email{andrewferguson@uchicago.edu}
\affiliation{%
  Pritzker School of Molecular Engineering, %
  The University of Chicago, %
  929 East 57th Street, Chicago, Illinois 60637, United States%
}

@@

We performed molecular dynamics simulations of four 10-base self-complementary double-stranded DNA sequences that we have previously studied by ultra-fast temperature-jump infrared spectroscopy \citep{Sanstead2016}: 5$^\prime$-ATATATATAT-3$^\prime$ (AT-all), 5$^\prime$-GATATATATC-3$^\prime$ (GC-end), 5$^\prime$-ATATGCATAT-3$^\prime$ (GC-core), and  5$^\prime$-ATGATATCAT-3$^\prime$ (GC-mix). We modeled the DNA sequences using the coarse-grained 3-Site-Per-Nucleotide v2 (3SPN.2) model that uses three spherical beads to represent the phosphate, deoxyribose sugar, and nitrogenous base of each nucleotide and employs anisotropic interaction potentials to accurately treat intra-strand base-stacking, inter-strand cross-stacking, and base pairing. \citep{Hinckley2013AnHybridization} The model was parameterized against experimental data on bond lengths, bend angles, torsional angles, base step energies, and base stacking free energies, and reliably reproduces the structure, melting temperatures, persistence lengths, and sequence, salt, concentration, and temperature effects on duplex formation. \citep{Hinckley2013AnHybridization} The model enables access to millisecond time scales and has been has widely-adopted to study numerous phenomenon including DNA packing in viral capsids, protein-DNA binding, and nucleosome unwrapping.\citep{Cordoba2017AIons, Lu2020OpenAWSEMSimulations, Lequieu2016Tension-dependentUnwrapping} Although the 3SPN.2 model was not directly parameterized against dynamical experimental data, we will show below that the predicted sequence-dependent kinetics and relaxation are, within a corrective scaling factor, in very good agreement with observed experimental trends.

@@

The details of the technique and processing used to acquire temperature-jump infrared (T-jump IR) data have been described previously.~\citep{Chung2007TransientKinetics,Jones2009Heterodyne-detectedSpectroscopy,Jones2012TransientInfrared} Briefly, heating was initiated through optical excitation of the O-D stretch overtone transition of \ce{D2O}. The 1.98 $\mu$m pulses (5 ns, 20 mJ, 20 Hz) used for heating were generated from the frequency-doubled output of a Nd:YAG (YG 980, Quantel) later sent through an optical parametric oscillator (OPO). Nonlinear IR spectra are collected from 5 ns to 50 ms delays after the T-jump. T-jump heterodyne-detected vibrational echo (t-HDVE) IR spectra were acquired with Fourier transform spectral interferometry,~\citep{Jones2009Heterodyne-detectedSpectroscopy} where the delay between the local oscillator (LO) and DVE signal was scanned in 5 fs steps between (-10) and 10 fs. t-HDVE spectra were acquired with a parallel pulse polarization scheme. Data are presented using the recovered T-jump dispersed pump-probe (t-DPP) spectrum. t-DPP data are reported as the difference spectra with the 49 or 50 ms spectrum, which corresponds the initial temperature spectrum.

The sample was placed between two 1 mm thick CaF2 windows separated by a 50 $\mu$m Teflon spacer enclosed in a home-built brass jacket. The initial temperature of the sample was set using a recirculating chiller connected to the brass sample jacket. The T-jump temperature change ($\Delta$T) was set to 14-16 $^\circ$C for all measurements and monitored using the change in transmission of the \ce{D2O} bend-libration combination band measured in the LO beam. The temperature change was quantified by comparing the change in transmission of the LO beam with a FTIR temperature series of \ce{D2O}.

@@

To computationally mimic the T-jump process in our 3SPN.2 simulations, we conducted 1 $\mu$s simulations of an initially hybridized DNA duplex over a range of temperatures and monitored its structural relaxation. We performed 120 independent simulations for each DNA sequence at each temperature, and from these extracted computational estimates of $k_d^\mathrm{slow}$ and fast $k_d^\mathrm{fast}$ (\blauw{Fig.~S4}). First, we tracked the slow response corresponding to duplex dissociation in our simulations by compiling the distribution of times at which both of the central base pairs first separate to a distance of 1.3 nm. This cutoff was selected as the distance beyond with the strands are effectively non-interacting and defines the dissociated state. We extracted our computational estimate of $k_d^\mathrm{slow}$ by fitting a decaying exponential to the fraction of hybridized sequences as a function of time $f_\mathrm{hybridized}(t) = \exp(-k_d^\mathrm{slow}t)$. Second, we tracked the fast response corresponding to terminal base pair fraying by compiling the distribution of times at which either of the terminal base pairs first separated to a distance of 1.3 nm, corresponding to a complete breakage of the WC interaction. We extracted our computational estimate of $k_d^\mathrm{fast}$ through a decaying exponential fit to the fraction of unfrayed sequences as a function of time $f_\mathrm{unfrayed}(t) = \exp(-k_d^\mathrm{fast}t)$.

We present in Fig.~\ref{fig:relaxation-comparison} a comparison of $k_d^\mathrm{slow}$ and $k_d^\mathrm{fast}$ estimated by computation and experiment. Although the 3SPN.2 model was not directly fitted against kinetic data,~\cite{Hinckley2013AnHybridization} its predictions of sequence dependent T-jump relaxation rates are, within a systematic scaling factor in time and systematic shift in temperature, in good agreement with observed experimental trends. It is well-known that the smoothing of the underlying free energy landscape induced by coarse-graining artificially accelerates the kinetics of coarse-grained molecular simulations and that different degrees of freedom may be accelerated by different factors.~\citep{Marrink2007TheSimulations, Fritz2011MultiscaleDynamics, Marrink2013PerspectiveModel} We find that the simulated slow responses corresponding to center-of-mass translation of the strands during dissociation of the duplex is $\sim$10$\times$ accelerated relative to experiment, whereas the fast responses corresponding to fraying of the terminal bases is $\sim$120$\times$ accelerated. We apply these sequence-independent scaling factors to our reported computational values. Although the 3SPN.2 model reproduces melting temperatures relatively well, we observed a systematic 4 K under-prediction in making the experimental comparison and so we apply a universal (+4) K corrective calibration to our computational results. We note that these empirical calibration factors to the 3SPN.2 predictions are applied only for the purposes of making an experimental comparison. The computational time scales and melting temperatures reported in the remainder of the paper are not corrected since we are only interested in the relative trends in the behaviors of the four sequences.

@@

\begin{figure}[ht!]
	\begin{center}
        \includegraphics[width=1.0\textwidth]{Fig2.pdf}
	\end{center}
\end{figure}
\begin{figure}[ht!]
        \caption{Thermodynamic and kinetic predictions of the sequence-specific MSMs for AT-all, GC-end, GC-core, and GC-mix. (a) \textbf{Macrostates.} Schematic representation of the seven metastable macrostates occupied by one or more of the four sequences: fully hybridized state (H), two-base out-of-register shifted states in the 5$^\prime$ (5S2) or 3$^\prime$ (3S2) direction, four-base out-of-register shifted states in the 5$^\prime$ (5S4) or 3$^\prime$ (3S4) direction, and the fully dissociated state (D). The line drawings represent the 10-base self-complementary sequences, where red-to-blue contacts indicates (possible) WC base pairing and black indicates an unbound bases. Adjacent to each line drawing we provide representative molecular structures corresponding to that macrostate. (b) \textbf{Thermodynamics.} Histograms reporting the number of the $10^7$ total frames within the 1 ms of simulation trajectories observed to occupy each of the seven macrostates, corresponding to our numerical estimates of the equilibrium occupancy probabilities. Uncertainties are calculated across 100 MSMs using a Bayesian MSM estimation are reported for each bar and are very small compared to the total counts. (c) \textbf{Kinetic networks.} MSMs illustrating the kinetic network for each sequence. The orange circles correspond to the macrostates occupied by each sequence and are labeled by the macrostate codes reported in panel a. The radii of the circles is proportional to the logarithm of the equilibrium occupancy probabilities reported in panel b. Molecular renderings of an illustrative snapshot from the coarse-grained molecular simulations are provided next to each macrostate. The grey arrows between macrostates indicate the presence of a probability flux between this pair of states at equilibrium and the arrow thickness is proportional to the flux. (To avoid congesting the diagram, arrows are not reported for probability fluxes lower than $3\times10^{-6}$.) The numerical value overlaid on each arrow reports the conditional probability that a system occupying the macrostate at the start of the arrow at time $t$ will transition to the macrostate at the end of the arrow by time $(t+\tau)$, where $\tau$=1.2 ns is the lag time corresponding to a single time step of the MSM. Large orange circles correspond to thermodynamically favorable states and large grey arrows correspond to kinetically favorable transitions. (d) \textbf{Kinetic time scales.} Distribution of MSM implied time scales for each sequence. The leading implied time scale corresponds to the characteristic time scale for hybridization/dehybridization and is approximately equal for all systems since simulations were conducted at the same concentration and at the respective melting temperatures. The higher order implied time scales correspond to a spectrum of kinetic relaxations between the constituent macrostates in the MSM corresponding to shifted and/or frayed states.
    }
    \label{fig:allseq_table}
\end{figure}

@@

We first compare the thermodynamic predictions for the equilibrium macrostate probabilities to those of predictive sequence-dependent models for the thermodynamics of DNA association. The nearest neighbor (NN) model is popular model of DNA hybridization thermodynamics that predicts the free energy of duplex formation as a sum over helix initiation terms and the hybridization free energies of nearest neighbor pairs of bases that account for both the specific WC pairings and the modulating effects of the local (i.e., nearest neighbor) environment.~\citep{SantaLucia1998AThermodynamics, Santalucia2004TheMotifs} The parameters of the NN model were estimated by regressing over 108 experimental measurements to furnish a predictive model for the free energy of association as a function of DNA sequence and explicitly account for stacking contributions of native pairs, internal mismatches, and dangling ends. We apply the NN model to estimate the (relative) free energy $F$ of each of the macrostates occupied by each of the four sequences. The free energy of each macrostate is related to its equilibrium occupancy probability $P$ via the statistical mechanical relationship $F = -k_B T \ln P + C$, where $T$ is temperature, $k_B$ is Boltzmann's constant, and $C$ is an additive constant reflecting our ignorance of the absolute scale of free energies. Since the additive constant cancels out in the evaluation of relative free energies between macrostates, this calculation provides a means to compare the free energies predicted by our MSMs (Fig.~\ref{fig:allseq_table}) to those of the NN model. A full accounting of the application of the NN model to the macrostates occupied by each of the four sequences is provided in the \blauw{Supporting Information}.

As illustrated in Fig.~\ref{fig:NN_table}, we see that the NN model tends to predict lower free energies for each macrostate compared to the MSM and therefore overestimate macrostate stability relative to the hybridized (H) state by as much as 14.7 kJ/mol. (The free energy of the hybridized state is set to zero by construction for both the MSM and NN models.) The only exception to this trend is the frayed state (F4) unique to CG-core, where the MSM and NN predictions are in good agreement. Although the trends are in qualitative agreement, what is the root of the quantitative discrepancy of the MSM and NN models in the 5S2, 3S2, 5S4, and 3S4 states? First, the MSM is constructed bottom-up from molecularly detailed 3SPN.2 simulations whereas the NN model is fitted top-down by regression against experimental data. There are approximations inherent in the 3SPN.2 model and the NN model was fitted to limited experimental data assuming a low-order expansion in terms of nearest neighbor additive contributions. Second, although 3SPN.2 is expected to capture some dangling end stabilization effects through base stacking and cross-stacking interactions, the model was not parameterized to fully capture the enthalpic contributions of the interaction of unbound bases with terminal base pairs, whereas this term is explicitly included within the NN model. Third, a well known deficiency of the NN model is the absence of any treatment of inert tails -- free bases that extend beyond the dangling end tend to destabilize the duplex.\citep{DiMichele2014EffectHybridization}

@@

We can further explore the role of inert tails upon macrostate stability by analyzing the AT-all and GC-end sequences that occupy out-of-register shifted macrostates 5S2 and 3S2 comprising a one-base inert tail and 5S4 and 3S4 comprising a three-base inert tail (Fig.~\ref{fig:allseq_table}). For both sequences, our simulations show that 5S2 is both the most stable of shifted states and has the smallest discrepancy (4.0 kJ/mol AT-all, 10.0 kJ/mol GC-end) compared to NN predictions (Fig.~\ref{fig:NN_table}b). The 3S2 and and 5S4 states have nearly the same deviation from NN predictions (7.0-7.2 kJ/mol AT-all, 14.3-14.7 kJ/mol GC-end), indicating that longer tails in the 5$^\prime$ direction are as destabilizing as shorter tails in the 3$^\prime$ direction. It is known from differential scanning calorimetry studies~\citep{Dickman2012ThermodynamicDNAs} that 3$^\prime$ inert tails are more destabilizing than 5$^\prime$ tails, with the differential behavior attributed to a combination of 5$^\prime$ tails preferentially stacking on the core duplex and 3$^\prime$ tails perturbing the duplex structure. \citep{Doktycz1990ThermodynamicC, Dickman2012ThermodynamicDNAs, DiMichele2014EffectHybridization}. Both the NN and MSM predictions for AT-all are consistent with this trend (i.e., $F_\mathrm{3S2} > F_\mathrm{5S2}$ and $F_\mathrm{3S4} > F_\mathrm{5S4}$). For GC-end, the MSM and NN models both predict the four shifted states to be less stable than the corresponding states in AT-all. This is in line with expectations since the terminal G:C pairs in GC-end decrease by two the number of available WC pairings in out-of-register shifted states compared to AT-all. The GC-end NN predictions run contrary to the expectation that the 3$^\prime$ inert tails should be more destabilizing than the 5$^\prime$ tails, whereas the MSM predictions are consistent with this trend. Indeed, the MSM model for GC-end does not identify the 3S4 macrostate as a stable metastable conformation for the duplex.

@@

\begin{figure}[ht!]
	\begin{center} 
        \includegraphics[width=160mm, scale=1]{Fig3.pdf}
        \caption{Comparison of the macrostate free energy predictions of the MSMs and nearest neighbor (NN) thermodynamic model.~\citep{SantaLucia1998AThermodynamics, Santalucia2004TheMotifs} (a) Free energies of each macrostate relative to the hybridized state that is assigned a free energy of zero. (b) Discrepancy between the macrostate free energy predictions of the MSM relative to the NN model. With the exception of the F4 frayed state, the NN model predicts lower free energies than the MSM predictions.}
        \label{fig:NN_table}
	\end{center}
\end{figure}

@@

The GC-core MSM represents a departure from the relatively rich and complex kinetic networks dominated by out-of-register shifted states to a much simpler one dominated by fraying (Fig.~\ref{fig:allseq_table}c). The MSM contains only three states -- hybridized H, dehybridized D, and frayed F4. The F4 state is unique to GC-core and contains up to six WC pairs -- the two central G:C core pairs and as many as four A:T pairs on one side or other of the core, while the other run of four A:T pairs remains free. (As expected by symmetry, the particular AT run that is made vs.\ free occurs with equal probability.) Our model reveals the absence of any direct hybridization or dehybridization transitions between the H and D states, with all pathways passing through the frayed state F4. Previous studies would suggest that hybridization of this sequence should proceed via a zippering mechanism, wherein upon formation of the strong central G:C WC base pairings the duplex helix should rapidly assemble in a middle-out fashion.~\citep{Romano2013DNADependence, Yin2011KineticsHybridization} Our results are partially consistent with this expectation, but reveal the frayed state F4 to be unexpectedly metastable, serving as a long-lived state with an mean life time of 1.8 ns. The stability of the state is attributable to the enthalpic stabilization offered by the up to six WC pairs and the entropic stabilization associated with the configurational entropy of the two free AT-tails.

@@

\section*{Supporting Information Available}

The following files are available free of charge.

\begin{itemize}
\item SI.pdf
\end{itemize}

>>>>

%\newcommand*{\rood}[1]{{\color{red}{#1}}}
%\newcommand*{\blauw}[1]{\textcolor{blue}{#1}}
\newcommand*{\blauw}[1]{{#1}}

%\newcommand*{\addref}[1]{\textcolor{red}{\{ADD REF: #1\}}}
\newcommand*{\noter}[1]{\textcolor{red}{[[#1]]}}		% notes 
%\newcommand*{\noter}[1]{}					% notes off
%\newcommand*{\noteb}[1]{\textcolor{blue}{[[#1]]}}		% notes

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Meta-data block
%% ---------------
%% Each author should be given as a separate \author command.
%%
%% Corresponding authors should have an e-mail given after the author
%% name as an \email command. Phone and fax numbers can be given
%% using \phone and \fax, respectively; this information is optional.
%%
%% The affiliation of authors is given after the authors; each
%% \affiliation command applies to all preceding authors not already
%% assigned an affiliation.
%%
%% The affiliation takes an option argument for the short name.  This
%% will typically be something like "University of Somewhere".
%%
%% The \altaffiliation macro should be used for new address, etc.
%% On the other hand, \alsoaffiliation is used on a per author basis
%% when authors are associated with multiple institutions.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\author{Michael S. Jones}
\affiliation{%
  Pritzker School of Molecular Engineering, %
  The University of Chicago, %
  5640 South Ellis Avenue, Chicago, Illinois 60637, United States%
}

@@

\author{Andrew L. Ferguson}
\email{andrewferguson@uchicago.edu}
\affiliation{%
  Pritzker School of Molecular Engineering, %
  The University of Chicago, %
  5640 South Ellis Avenue, Chicago, Illinois 60637, United States%
}

@@

We performed molecular dynamics simulations of four 10-base self-complementary double-stranded DNA sequences that we have previously studied by temperature-jump infrared spectroscopy \citep{Sanstead2016}: 5$^\prime$-ATATATATAT-3$^\prime$ (AT-all), 5$^\prime$-GATATATATC-3$^\prime$ (GC-end), 5$^\prime$-ATATGCATAT-3$^\prime$ (GC-core), and  5$^\prime$-ATGATATCAT-3$^\prime$ (GC-mix). We modeled the DNA sequences using the coarse-grained 3-Site-Per-Nucleotide v2 (3SPN.2) model that uses three spherical beads to represent the phosphate, deoxyribose sugar, and nitrogenous base of each nucleotide and employs anisotropic interaction potentials to accurately treat intra-strand base-stacking, inter-strand cross-stacking, and base pairing. \citep{Hinckley2013AnHybridization} The model was parameterized against experimental data on bond lengths, bend angles, torsional angles, base step energies, and base stacking free energies, and reliably reproduces the structure, melting temperatures, persistence lengths, and sequence, salt, concentration, and temperature effects on duplex formation. \citep{Hinckley2013AnHybridization} The model enables access to millisecond time scales and has been has widely-adopted to study numerous phenomenon including DNA packing in viral capsids, protein-DNA binding, and nucleosome unwrapping.\citep{Cordoba2017AIons, Lu2020OpenAWSEMSimulations, Lequieu2016Tension-dependentUnwrapping} Although the 3SPN.2 model was not directly parameterized against dynamical experimental data, we will show below that the predicted sequence-dependent kinetics and relaxation are, within a corrective scaling factor, in very good agreement with observed experimental trends.

@@

The details of the technique and processing used to acquire temperature-jump infrared (T-jump IR) data have been described previously.~\citep{Chung2007TransientKinetics,Jones2009Heterodyne-detectedSpectroscopy,Jones2012TransientInfrared} Briefly, heating was initiated through optical excitation of the O-D stretch overtone transition of \ce{D2O}. The 1.98 $\mu$m pulses (5 ns, 20 mJ, 20 Hz) used for heating were generated from the frequency-doubled output of a Nd:YAG laser sent through an optical parametric oscillator. Nonlinear IR spectra are collected from 5 ns to 50 ms delays after the T-jump with a synchronized 1 kHz spectrometer. T-jump heterodyne-detected vibrational echo (t-HDVE) IR spectra were acquired with Fourier transform spectral interferometry,~\citep{Jones2009Heterodyne-detectedSpectroscopy} where the delay between the local oscillator (LO) and DVE signal was scanned in 5 fs steps between (-10) and 10 fs. t-HDVE spectra were acquired with a parallel pulse polarization scheme and presented as a dispersed pump-probe (t-DPP) spectrum. t-DPP data are reported as the difference spectra relative to the initial temperature.

The sample was placed between two 1 mm thick CaF2 windows separated by a 50 $\mu$m Teflon spacer enclosed in a brass jacket. The initial temperature of the sample was set using a recirculating chiller connected to the brass sample jacket. The T-jump temperature change ($\Delta$T) was set to 14-16 $^\circ$C for all measurements and monitored using the change in transmission of the \ce{D2O} bend-libration combination band measured in the LO beam. The temperature change was quantified by comparing the change in transmission of the LO beam with a FTIR temperature series of \ce{D2O}.

@@

To computationally mimic the T-jump process in our 3SPN.2 simulations, we conducted 1 $\mu$s simulations of an initially hybridized DNA duplex over a range of temperatures and monitored its structural relaxation. We performed 120 independent simulations for each DNA sequence at each temperature, and from these extracted computational estimates of $k_d^\mathrm{slow}$ and fast $k_d^\mathrm{fast}$ (\blauw{Fig.~S4}). First, we tracked the slow response corresponding to duplex dissociation in our simulations by compiling the distribution of times at which both of the central base pairs first separate to a distance of 1.3 nm starting from an initial fully hybridized duplex. This cutoff was selected as the distance beyond with the strands are effectively non-interacting and defines the dissociated state. We extracted our computational estimate of $k_d^\mathrm{slow}$ by fitting a decaying exponential to the fraction of hybridized sequences as a function of time $f_\mathrm{hybridized}(t) = \exp(-k_d^\mathrm{slow}t)$. Second, we tracked the fast response corresponding to terminal base pair fraying by compiling the distribution of times at which either of the terminal base pairs first separated to a distance of 1.3 nm, corresponding to a complete breakage of the WC interaction. We extracted our computational estimate of $k_d^\mathrm{fast}$ through a decaying exponential fit to the fraction of unfrayed sequences as a function of time $f_\mathrm{unfrayed}(t) = \exp(-k_d^\mathrm{fast}t)$.

We present in Fig.~\ref{fig:relaxation-comparison} a comparison of $k_d^\mathrm{slow}$ and $k_d^\mathrm{fast}$ estimated by computation and experiment. Although the 3SPN.2 model was not directly fitted against kinetic data,~\cite{Hinckley2013AnHybridization} its predictions of sequence dependent T-jump relaxation rates are, within a systematic scaling factor in time and systematic shift in temperature, in good agreement with observed experimental trends. It is well-known that the smoothing of the underlying free energy landscape induced by coarse-graining artificially accelerates the kinetics of coarse-grained molecular simulations and that different degrees of freedom may be accelerated by different factors.~\citep{Marrink2007TheSimulations, Fritz2011MultiscaleDynamics, Marrink2013PerspectiveModel} We find that the simulated slow responses corresponding to center-of-mass translation of the strands during dissociation of the duplex is $\sim$10$\times$ accelerated relative to experiment, whereas the fast responses corresponding to fraying of the terminal bases is $\sim$120$\times$ accelerated. We apply these sequence-independent scaling factors to our reported computational values. Although the 3SPN.2 model reproduces melting temperatures relatively well, we observed a systematic 4 K under-prediction in making the experimental comparison and so we apply a universal (+4) K corrective calibration to our computational results. We note that these empirical calibration factors to the 3SPN.2 predictions are applied only for the purposes of making an experimental comparison, but note that there are uncertainties introduced by assuming that the equilibrium dynamics at fixed temperature can be compared directly to relaxation kinetics following a 15$^\circ$C T-jump. The computational time scales and melting temperatures reported in the remainder of the paper are not corrected since we are only interested in the relative trends in the behaviors of the four sequences.

@@

\begin{figure}[ht!]
	\begin{center}
        \includegraphics[width=1.0\textwidth]{Fig2.pdf}
	\end{center}
\end{figure}
\begin{figure}[ht!]
        \caption{Thermodynamic and kinetic predictions of the sequence-specific MSMs for AT-all, GC-end, GC-core, and GC-mix. (a) \textbf{Macrostates.} Schematic representation of the seven metastable macrostates occupied by one or more of the four sequences: fully hybridized state (H), two-base out-of-register shifted states in the 5$^\prime$ (5S2) or 3$^\prime$ (3S2) direction, four-base out-of-register shifted states in the 5$^\prime$ (5S4) or 3$^\prime$ (3S4) direction, and the fully dissociated state (D). The line drawings represent the 10-base self-complementary sequences, where red-to-blue contacts indicates (possible) WC base pairing and black indicates an unbound bases. Adjacent to each line drawing we provide representative molecular structures corresponding to that macrostate. (b) \textbf{Thermodynamics.} Histograms reporting the number of the $10^7$ total frames within the 1 ms of simulation trajectories observed to occupy each of the seven macrostates, corresponding to our numerical estimates of the equilibrium occupancy probabilities. Uncertainties are calculated across 100 MSMs using a Bayesian MSM estimation are reported for each bar and are very small compared to the total counts. \noter{Add percentages to blue bars in white text.} (c) \textbf{Kinetic networks.} MSMs illustrating the kinetic network for each sequence. The orange circles correspond to the macrostates occupied by each sequence and are labeled by the macrostate codes reported in panel a. The radii of the circles is proportional to the logarithm of the equilibrium occupancy probabilities reported in panel b. Molecular renderings of an illustrative snapshot from the coarse-grained molecular simulations are provided next to each macrostate. The grey arrows between macrostates indicate the presence of a probability flux between this pair of states at equilibrium and the arrow thickness is proportional to the flux. (To avoid congesting the diagram, arrows are not reported for probability fluxes lower than $3\times10^{-6}$.) The numerical value overlaid on each arrow reports the conditional probability that a system occupying the macrostate at the start of the arrow at time $t$ will transition to the macrostate at the end of the arrow by time $(t+\tau)$, where $\tau$=1.2 ns is the lag time corresponding to a single time step of the MSM. Large orange circles correspond to thermodynamically favorable states and large grey arrows correspond to kinetically favorable transitions. (d) \textbf{Kinetic time scales.} Distribution of MSM implied time scales for each sequence. The leading implied time scale corresponds to the characteristic time scale for hybridization/dehybridization and is approximately equal for all systems since simulations were conducted at the same concentration and at the respective melting temperatures. The higher order implied time scales correspond to a spectrum of kinetic relaxations between the constituent macrostates in the MSM corresponding to shifted and/or frayed states.
    }
    \label{fig:allseq_table}
\end{figure}

@@

We first compare the thermodynamic predictions for the equilibrium macrostate probabilities to those of predictive sequence-dependent models for the thermodynamics of DNA association. The nearest neighbor (NN) model is popular model of DNA hybridization thermodynamics that predicts the free energy of duplex formation as a sum over helix initiation terms and the hybridization free energies of nearest neighbor pairs of bases that account for both the specific WC pairings and the modulating effects of the local (i.e., nearest neighbor) environment.~\citep{SantaLucia1998AThermodynamics, Santalucia2004TheMotifs} The parameters of the NN model were estimated by regressing over 108 experimental measurements to furnish a predictive model for the free energy of association as a function of DNA sequence and explicitly account for stacking contributions of native pairs, internal mismatches, and dangling ends. We apply the NN model to estimate the (relative) free energy $F$ of each of the macrostates occupied by each of the four sequences. The free energy of each macrostate is related to its equilibrium occupancy probability $P$ via the statistical mechanical relationship $F = -k_B T \ln P + C$, where $T$ is temperature, $k_B$ is Boltzmann's constant, and $C$ is an additive constant reflecting our ignorance of the absolute scale of free energies. Since the additive constant cancels out in the evaluation of relative free energies between macrostates, this calculation provides a means to compare free energy differences predicted by our MSMs (Fig.~\ref{fig:allseq_table}) to those of the NN model. For clarity of presentation, we specify the additive constant $C$ in both the MSM and NN models by arbitrarily assigning the free energy of the hybridized state H to define the zero of free energy. A full accounting of the application of the NN model to the macrostates occupied by each of the four sequences is provided in the \blauw{Supporting Information}.

As illustrated in Fig.~\ref{fig:NN_table}, we see that the NN model tends to predict lower free energies for each macrostate compared to the MSM and therefore overestimate macrostate stability relative to the hybridized state by as much as 14.7 kJ/mol. The only exception to this trend is the frayed state (F4) unique to CG-core, where the MSM and NN predictions are in good agreement. Although the trends are in qualitative agreement, what is the root of the quantitative discrepancy of the MSM and NN models in the predicted relative stability of the 5S2, 3S2, 5S4, and 3S4 states? First, the MSM is constructed bottom-up from molecularly detailed 3SPN.2 simulations whereas the NN model is fitted top-down by regression against experimental data. There are approximations inherent in the 3SPN.2 model, not least of which is the coarse-grained representation that integrates over atomic degrees of freedom, and in the NN model that was fitted to limited experimental data assuming a low-order expansion in terms of nearest neighbor additive contributions. Second, although 3SPN.2 is expected to capture some dangling end stabilization effects through base stacking and cross-stacking interactions, the model was not parameterized to fully capture the enthalpic contributions of the interaction of unbound bases with terminal base pairs, whereas this term is explicitly included within the NN model. Third, a well known deficiency of the NN model is the absence of any treatment of inert tails -- free bases that extend beyond the dangling end tend to destabilize the duplex.\citep{DiMichele2014EffectHybridization}

@@

We can further explore the role of inert tails upon macrostate stability by analyzing the AT-all and GC-end sequences that occupy out-of-register shifted macrostates 5S2 and 3S2 comprising a one-base inert tail and 5S4 and 3S4 comprising a three-base inert tail (Fig.~\ref{fig:allseq_table}). For both sequences, our simulations show that 5S2 is both the most stable of shifted states relative to H (Fig.~\ref{fig:NN_table}a) and has the smallest discrepancy (4.0 kJ/mol AT-all, 10.0 kJ/mol GC-end) compared to NN predictions (Fig.~\ref{fig:NN_table}b). The 3S2 and and 5S4 states have nearly the same deviation from NN predictions (7.0-7.2 kJ/mol AT-all, 14.3-14.7 kJ/mol GC-end), indicating that longer tails in the 5$^\prime$ direction are as destabilizing as shorter tails in the 3$^\prime$ direction. It is known from differential scanning calorimetry studies~\citep{Dickman2012ThermodynamicDNAs} that 3$^\prime$ inert tails are more destabilizing than 5$^\prime$ tails, with the differential behavior attributed to a combination of 5$^\prime$ tails preferentially stacking on the core duplex and 3$^\prime$ tails perturbing the duplex structure. \citep{Doktycz1990ThermodynamicC, Dickman2012ThermodynamicDNAs, DiMichele2014EffectHybridization}. Both the NN and MSM predictions for AT-all are consistent with this trend (i.e., $F_\mathrm{3S2} > F_\mathrm{5S2}$ and $F_\mathrm{3S4} > F_\mathrm{5S4}$). For GC-end, the MSM and NN models both predict the four shifted states to be less stable relative to the hybridized state than the corresponding predictions for AT-all. This is in line with expectations since the terminal G:C pairs in GC-end decrease by two the number of available WC pairings in out-of-register shifted states compared to AT-all. The GC-end NN predictions run contrary to the expectation that the 3$^\prime$ inert tails should be more destabilizing than the 5$^\prime$ tails, whereas the MSM predictions are consistent with this trend. Indeed, the MSM model for GC-end does not identify the 3S4 macrostate as a stable metastable conformation for the duplex.

@@

\begin{figure}[ht!]
	\begin{center} 
        \includegraphics[width=160mm, scale=1]{Fig3.pdf}
        \caption{Comparison of the macrostate free energy predictions of the MSMs and nearest neighbor (NN) thermodynamic model.~\citep{SantaLucia1998AThermodynamics, Santalucia2004TheMotifs} (a) Free energies of each macrostate relative to the hybridized state that is assigned a free energy of zero. (b) Discrepancy between the macrostate free energy predictions of the MSM relative to the NN model. With the exception of the F4 frayed state, the NN model predicts lower free energies than the MSM predictions. \noter{(i) y-axis label in a: ``free energy relative to hybridized state, F (kJ/mol)'' (ii) y-axis label in b: ``free energy discrepancy between MSM and NN model relative to hybridized state, $\Delta$F (kJ/mol)'' -- might have to go on two lines}}
        \label{fig:NN_table}
	\end{center}
\end{figure}

@@

The GC-core MSM represents a departure from the relatively rich and complex kinetic networks dominated by out-of-register shifted states to a much simpler one dominated by fraying (Fig.~\ref{fig:allseq_table}c). The MSM contains only three states -- hybridized H, dehybridized D, and frayed F4. The F4 state is unique to GC-core and contains up to six WC pairs -- the two central G:C core pairs and as many as four A:T pairs on one side or other of the core, while the other run of four A:T pairs remains free. (As expected by symmetry, the particular AT run that is made vs.\ free occurs with equal probability.) \noter{Maybe add a sentence here on other frayed states containing 3, 2, 1 frayed bases but not metastable within our MSM.} Our model reveals the absence of any direct hybridization or dehybridization transitions between the H and D states, with all pathways passing through the frayed state F4. Previous studies would suggest that hybridization of this sequence should proceed via a zippering mechanism, wherein upon formation of the strong central G:C WC base pairings the duplex helix should rapidly assemble in a middle-out fashion.~\citep{Romano2013DNADependence, Yin2011KineticsHybridization} Our results are partially consistent with this expectation, but reveal the frayed state F4 to be unexpectedly metastable, serving as a long-lived state with an mean life time of 1.8 ns. The stability of the state is attributable to the enthalpic stabilization offered by the up to six WC pairs and the entropic stabilization associated with the configurational entropy of the two free AT-tails.

@@

\section*{Supporting Information}

Details of nearest neighbor (NN) model calculations, supplementary figures illustrating selection of optimal number of SRV slow modes for each sequence, convergence of MSM implied time scales, MSM Chapman-Kolmogorov tests, ``computational T-jump'' calculations, and GC-mix hybridization and dehybridization trajectories.
